%  ========================================================================
%  Copyright (c) 1985-2014 The University of Washington
%
%  Licensed under the Apache License, Version 2.0 (the "License");
%  you may not use this file except in compliance with the License.
%  You may obtain a copy of the License at
%
%      http://www.apache.org/licenses/LICENSE-2.0
%
%  Unless required by applicable law or agreed to in writing, software
%  distributed under the License is distributed on an "AS IS" BASIS,
%  WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
%  See the License for the specific language governing permissions and
%  limitations under the License.
%  ========================================================================
%

% Documentation for University of Washington thesis LaTeX document class
% by Jim Fox
% fox@washington.edu
%
%    Revised for version 2015/03/03 of uwthesis.cls
%
%    This document is contained in a single file ONLY because
%    I wanted to be able to distribute it easily.  A real thesis ought
%    to be contained on many files (e.g., one for each chapter, at least).
%
%    To help you identify the files and sections in this large file
%    I use the string '==========' to identify new files.
%
%    To help you ignore the unusual things I do with this sample document
%    I try to use the notation
%       
%    % --- sample stuff only -----
%    special stuff for my document, but you don't need it in your thesis
%    % --- end-of-sample-stuff ---


%    Printed in twoside style now that that's allowed
%
 
\documentclass [11pt, proquest] {uwthesis}[2015/03/03]
 
%
% The following line would print the thesis in a postscript font 

\usepackage{natbib}
\def\bibpreamble{\protect\addcontentsline{toc}{chapter}{Bibliography}}

\setcounter{tocdepth}{1}  % Print the chapter and sections to the toc
 

% ==========   Local defs and mods
%
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsthm}

% BRH defs
\usepackage{longtable}
\usepackage{booktabs}
\def\tightlist{}

\begin{document}
 
% ==========   Preliminary pages
%
% ( revised 2012 for electronic submission )
%

\prelimpages
 
%
% ----- copyright and title pages
%
\Title{Reducing errors in simulated satellite views of clouds from large-scale models}
\Author{Benjamin R. Hillman}
\Year{2016}
\Program{Department of Atmospheric Sciences}

\Chair{Thomas P. Ackerman}{}{}
\Signature{Roger T. Marchand}
\Signature{Robert Wood} 
\Signature{Daehyun Kim}

\copyrightpage

\titlepage  

%
% ----- abstract
%

\setcounter{page}{-1}
\abstract{%
[TODO: fix this...this is just a placeholder] The representation of clouds in global climate models (GCMs) is recognized as a primary source of uncertainty in projections of future climate change scenarios and in equilibrium climate sensitivity. A fundamental test of the representation of clouds in models is evaluating the simulation of present-day climate, for which we have available observations. Satellite retrievals of cloud properties provide an attractive baseline for this evaluation, because they provide near global coverage, and oftentimes long records (decades or more). However, comparisons of modeled and satellite-retrieved cloud properties are challenged by fundamental differences between what a model can represent, and what can be observed from space. Satellite simulators have emerged in recent decades in an attempt to account for these differences by producing psuedo-retrievals of cloud properties from the model diagnosed description of the atmosphere. These simulators have been the subject of a number of evaluation studies, but a lack of rigorous treatment of the uncertainties in this framework undermines these comparisons. In this thesis, potential sources of uncertainty in simulating retrievals from modeled cloud descriptions are evaluated and quantified. It is shown that assumptions about the structure and variability at scales below which can be resolved with current global models can significantly affect the simulated retrievals, and an improved framework for downscaling cloud properties from the large-scale model to the satellite pixel scale is shown to reduce these errors.
}
 
%
% ----- contents & etc.
%
\tableofcontents
\listoffigures
\listoftables  % I have no tables
 
%
% ----- glossary 
%
%\chapter*{Glossary}      % starred form omits the `chapter x'
%\addcontentsline{toc}{chapter}{Glossary} 
%\thispagestyle{plain}
%
%\input{glossary}
 
%
% ----- acknowledgments
%
\acknowledgments{% \vskip2pc
  % {\narrower\noindent
  TODO: put some acknowledgments here.
  % \par}
}

%
% ----- dedication
%
\dedication{\begin{center}TODO: put a dedication here.\end{center}}

%
% end of the preliminary pages
 
%
% ==========      Text pages
%

\textpages
 
% ========== Input chapters
\input{latex/01_intro}
\input{latex/02_misr}
\input{latex/03_subgrid1}
\input{latex/04_subgrid2}
\input{latex/05_cmip5}

%
% ==========   Bibliography
%
%\nocite{*}   % include everything in the uwthesis.bib file
\bibliographystyle{ametsoc}
\bibliography{dissertation}
%
% ==========   Appendices
%
%\appendix
%\raggedbottom\sloppy
% 
%% ========== Appendix A
% 
%\chapter{A useful appendix}
% 
%Here we have a useful description of something too involved to put in the text itself. Or maybe a link to a jupyter notebook containing analysis. Or something else, who knows.

\vita{TODO: add short bio.}
%\vita{Jim Fox is a Software Engineer with UW Information Technology at the University of Washington.
%His duties do not include maintaining this package.  That is rather
%an avocation which he enjoys as time and circumstance allow.

%He welcomes your comments to {\tt fox@uw.edu}.
%}

\end{document}
